Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-14 csf-2017/paper.tex
Some slight changes in the text about macro-instructions.

\textbf{Reusable macro instructions} With the high-level approach in
mind, we define the following macro instructions: 
\texttt{fetch $r$ $\var{name}$} fetches the capability related to
$\var{name}$ from the linking table and stores it in register
$r$. 
\texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
macros that uses the stack assumes a stack pointer in
register $r_\stk$. 
\texttt{prepstk $r$} prepares the stack by checking it has permission
$\rwlx$ and making it conform to our stack discipline. 
\texttt{scall  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}
Jumps to the capability in register $r$ in a safe manner. That is it
pushes local state, stack pointer and in code return pointer to
the stack, sets up the necessary on stack return code and creates a
return pointer, restricts stack pointer, clears the unused part
of the stack, clears the necessary registers and jumpsto $r$. Upon
return, the private state is restored.
\texttt{mclear $r$} clears all the memory the capability in register
$r$ has authority over.
\texttt{rclear $\var{regSet}$} clears all the registers in $\var{regSet}$
\texttt{reqglob $r$} checks whether the word in register $r$ is a
global capability.
\texttt{crtcls $\overline{(x_i,r_i)}$ $r$} allocates a closure pair
where the capability in register $r$ is used to point to the program
and a new environment is allocated where the contents of $\overline{r_i}$ is
stored. In the program referred to by $r$, an implicit fetch
happens when an instruction refers to $x_i$.

@@

\subsection{Worlds}
%\todo[inline]{Find a good place to describe monotonicity of
%interpretation function.}
% \lau{this may not be an important enough point to include it.}
\begin{figure}[htbp]
  \centering
  \includegraphics{w11}
  \caption{Illustration of the relation between local and global capabilities and temporary and permanent regions. The table represents a memory and the colored fields are regions governing parts of the memory. Roughly speaking, global capabilities should never depend on memory governed by temporary regions.}
  \label{fig:cap-world}
\end{figure}
A world is a finite map from region names, modeled as natural numbers, to regions.
We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
%$\perma$, $\temp$, and $\revoked$. The
%$\perma$ and $\temp$ 
Each permanent and temporary region contains a state transition system, with
public and private transitions, which describes the protocol governing
the evolution of the memory described by the region, similarly to what
has been used in logical relations for high-level
languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. 
The protocols imposed on memory by permanent regions stay in place
indefinitely. This means that any capability, local or global, can
depend on these protocols. The protocols imposed on memory by temporary
regions can be revoked at certain points during execution,
corresponding to the points when local capabilities can be
revoked. This means that local capabilities safely can depend on the
protocols imposed on memory by temporary regions, whereas global
capabilities cannot, since a global capability may stay around
throughout the execution of a program, but a temporary region may be
revoked. This is illustrated in \figurename~\ref{fig:cap-world}.

@@

We believe local capabilities

>>>>

\textbf{Reusable macro instructions} With the high-level approach in mind, we
define the following macro instructions: \texttt{fetch $r$ $\var{name}$} fetches
the capability related to $\var{name}$ from the linking table and stores it in
register $r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
macros that uses the stack assumes a stack pointer in register $r_\stk$.
\texttt{prepstk $r$} is used when a callback is invoked by the adversary and
prepares the received stack pointer by checking that it has permission $\rwlx$
and making it conform to our stack discipline. \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps to the
capability in register $r$ in the manner described above. That is, it pushes local state (the
contents of registers $\overline{r_{\var{priv}}}$), stack pointer and return
pointer to the stack, sets up the necessary on stack return code and creates a
return pointer, restricts stack pointer, clears the unused part of the stack,
clears the necessary registers and jumps to $r$. Upon return, the private state
is restored. \texttt{mclear $r$} clears all the memory the capability in
register $r$ has authority over. \texttt{rclear $\var{regSet}$} clears all the
registers in $\var{regSet}$ \texttt{reqglob $r$} checks whether the word in
register $r$ is a global capability. \texttt{crtcls $\overline{(x_i,r_i)}$ $r$}
allocates a closure pair where the capability in register $r$ is used to point
to the program and a new environment is allocated where the contents of
$\overline{r_i}$ is stored. In the program referred to by $r$, an implicit fetch
happens when an instruction refers to $x_i$.

@@

\subsection{Worlds}
%\todo[inline]{Find a good place to describe monotonicity of
%interpretation function.}
% \lau{this may not be an important enough point to include it.}
\begin{figure}
  \centering
  \includegraphics{w11}
  \caption{Illustration of the relation between local and global capabilities and temporary and permanent regions. The table represents a memory and the colored fields are regions governing parts of the memory. Roughly speaking, global capabilities should never depend on memory governed by temporary regions.}
  \label{fig:cap-world}
\end{figure}
A world is a finite map from region names, modeled as natural numbers, to regions.
We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
%$\perma$, $\temp$, and $\revoked$. The
%$\perma$ and $\temp$ 
Each permanent and temporary region contains a state transition system, with
public and private transitions, which describes the protocol governing
the evolution of the memory described by the region, similarly to what
has been used in logical relations for high-level
languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. 
The protocols imposed on memory by permanent regions stay in place
indefinitely. This means that any capability, local or global, can
depend on these protocols. The protocols imposed on memory by temporary
regions can be revoked at certain points during execution,
corresponding to the points when local capabilities can be
revoked. This means that local capabilities safely can depend on the
protocols imposed on memory by temporary regions, whereas global
capabilities cannot, since a global capability may stay around
throughout the execution of a program, but a temporary region may be
revoked. This is illustrated in \figurename~\ref{fig:cap-world}.

@@
